---
title: "project-1"
author: "harsh"
date: "5/20/2017"
output: html_document
---

Question 1 - Import File ???bank-full.csv??? , ensure that categorical variables are imported as characters , not factors.
Take a quick look at type of variables in the data.

Import Project file
Answer Question NO 1 (Q-1)
Import File "bank-full.csv" 

```{r}

d = read.csv('bank-full.csv',sep= ";" ,stringsAsFactors = FALSE)
str(d)
library(dplyr)
glimpse(d)
head(d)

```

```{r}
head(d,20)
tail(d,20)
```

```{r}
library(psych)
summary(d)
describe(d)
```


Question no 2:- Find out mean ,std deviations, q1 , q3 and IQR for the variables age and balance. Check if they are
following normal distribution. If they are following normal distribution then calculate outlier limits as :
mean ?? 3*sd, otherwise calculate outlier limits as [q1 - 1.5 * IQR, q3 + 1.5 * IQR]. Find out which
observations exceed these limits, remove them from the data.


Answer for question for 2
#Find out mean ,std deviations, q1 , q3 and IQR for the variables age and balance. 

```{r}
library(stats)
sapply(d[,c(1,6)],summary, header= TRUE)   # check Mean, q1 , q3
```

```{r}

sapply(d[,c(1,6)],sd) # std deviations

```


```{r}

sapply(d[c(1,6)],IQR) # IQR

```
Check they are following normal distribution.

first we check hypothesis testing
```{r}
t.test(d$age,d$balance)
var.test(d$age,d$balance)
t.test(d$age,d$balance,var.equal = TRUE)
```

Normality Test

```{r}
library(nortest)
x <- d$age
y <- d$balance
ad.test(x)
ad.test(y)
```

Normal Quantile-Quantile plot for age
```{r}
qqnorm(x)
qqline(x, col = "red")
```
Normal Quantile-Quantile plot for balance


```{r}
qqnorm(y)
qqline(y, col = "red")
```


Note :  large portion of the data points do not fall on the straight line

----------------------
Visual Normality Test
--------------------
```{r}
library(ggplot2)
ab=data.frame(x,y)
ggplot(ab,aes(x))+geom_density(color="red")+ stat_function(fun=dnorm,aes(x),color="green")+ ggtitle("Visual Normality Test for Age ")
```


```{r}
ggplot(ab,aes(y))+geom_density(color="red")+ stat_function(fun=dnorm,aes(y),color="green")+ ggtitle("Visual Normality Test for Balance ")
```


```{r}
quantile(d$age)
age1 <- quantile(d$age)[3]-quantile(d$age)[2]
age2 <- quantile(d$age)[4]-quantile(d$age)[3]
print(age2-age1)
print(age1==age2)
```
```{r}


bal1 <- quantile(d$balance)[3]-quantile(d$balance)[2]
bal2 <- quantile(d$balance)[4]-quantile(d$balance)[3]
print(bal2-bal1)
print(bal1==bal2)

```

calculate outlier limits as :
mean ?? 3*sd, otherwise calculate outlier limits as [q1 - 1.5 * IQR, q3 + 1.5 * IQR].


Plot histogram for variables age
```{r}
hist(d$age)
```
Plot histogram for variables alance
```{r}
hist(d$balance)
```


calculate  outlier limits for age

```{r}
c=1.5
age = quantile(d$age)
bal = quantile(d$balance)
print("lower limit age")
out.age1=age[2] - c * IQR(d$age) 
out.age1
print("upper limit age")
out.age2 <- age[4] + c * IQR(d$age)
out.age2


```
calculate  numbers of outliers  in age var
```{r}
print("number of outliers in age")
sum(d$age <out.age1 | d$age>out.age2)
outlier_value <- boxplot.stats(d$age)$out
outlier_value
```
calculate  outlier limits for balance

```{r}
print("lower limit balance")
out.bal1 <- bal[2] - c * IQR(d$balance)
out.bal1
print("upper limit balance")
out.bal2 <- bal[4] + c * IQR(d$balance)
out.bal2
```
calculate  numbers of outliers  in balance var

```{r}
sum(d$balance <out.bal1 | d$balance>out.bal2)
outlier_value1 <- boxplot.stats(d$balance)$out
outlier_value1
```
```{r}
print("boxplot for age with outliers")
boxplot(d$age)
```

```{r}
print("boxplot for age without outliers")
boxplot(d$age,outline = FALSE)
```

```{r}
print("boxplot for balance with outliers")
boxplot(d$balance)
```

```{r}
print("boxplot for balance without outliers")
boxplot(d$balance,outline = FALSE)
```
```{r}
outlierReplace = function(dataframe, cols, rows, newValue = NA) {
   if (any(rows)) { 
           set(dataframe, rows, cols, newValue) }}
outlierReplace(d, "age", which(d$age > out.age2), NA)
outlierReplace(d, "age", which(d$age < out.age1), NA)
outlierReplace(d, "balance", which(d$balance > out.bal2), NA)
outlierReplace(d, "balance", which(d$balance < out.bal2), NA)
describe(d)
str(d)
```

```{r}
#Ans for question 3
#cross table for job and y
t1=table(d$job, d$y)

p1=prop.table(t1,1)

round(p1,2)
str(d)
```


```{r}
#cross table for month  and y
t2=table(d$month, d$y)
p2=prop.table(t2,1)
round(p2,2)
```


```{r}
#clubbing of categories
library(dplyr)
set.seed(123)
dnew=d%>%
       mutate( job.j1=as.numeric(job %in% c("admin.","self-employed","unknown")),
job.j2=as.numeric(job %in% c("housemaid","services")),
blue.collar=as.numeric(job == "blue-collar"),
entrep=as.numeric(job == "entrepreneur"),
mngmt=as.numeric(job == "management"),
ret=as.numeric(job == "retired"),
tech  = as.numeric(job=="technician"),
unemp=as.numeric(job == "unemployed"))%>%
        select(-job)
str(dnew)

#creating new catgorical var for job 
cate.job <- data.frame(dnew[,17:24])
glimpse(cate.job)

```


```{r}
str(dnew)[,24:33]
#clubbing of categories
set.seed(123)
dnew=dnew%>%
       mutate( month.m1=as.numeric(month %in% c("jun","jan","nov")),
        aug1=as.numeric(month == "aug"),
        july=as.numeric(month == "jul"),
        feb1=as.numeric(month == "feb"),
        may1=as.numeric(month == "may"),
        apr1  = as.numeric(month=="apr"),
        dec1=as.numeric(month == "dec"),
        sep1=as.numeric(month == "sep"),
        oct1=as.numeric(month == "oct"))%>%
        select(-month)
str(dnew)

#creating new catgorical var for month 
cate.month <- data.frame(dnew[,24:32])

glimpse(cate.month)
```
QUESTION 4

Write a
function which takes a dataset and categorical variable names as input and returns dataset with dummy
variables for that categorical variable.
```{r}

Dummies1=function(df,dvar){
  
  t=table(df[,dvar])
  
  t=sort(t)
  
  l2=names(t)[-1]
  
  for(l1 in l2){
    name=paste(dvar,l1,sep=".")
    name=gsub(" ","",name)
    name=gsub("-","",name)
    name=gsub("\\?","Q",name)
    
    df[,name]=as.numeric(df[,dvar]==l1)
  }
  
  df[,dvar]=NULL
  return(df)
}

##### 

#df = Dummies1(df,"dvar")

```

Question 5  Create pi-chart for the variable education with response y as fill.

```{r}
library(ggplot2)
ggplot(dnew, aes(x = "", fill = education)) + geom_bar() + coord_polar(theta = "y")
```


Question 6-Using function above [written in Q4] or otherwise create dummy variables for the categorical variables


```{r}
mydata <- d
mydata <- Dummies1(mydata,"job")
mydata <- Dummies1(mydata,"marital")
mydata <- Dummies1(mydata,"education")
mydata <- Dummies1(mydata,"default")
mydata <- Dummies1(mydata,"housing")
mydata <- Dummies1(mydata,"loan")
mydata <- Dummies1(mydata,"contact")
mydata <- Dummies1(mydata,"month")
mydata <- Dummies1(mydata,"poutcome")
mydata <- Dummies1(mydata,"y")
str(mydata)
```





